---
title: 'ML #3 Predictive Policing'
author: "Prof. Fichman & Prof. Harris"
date: "10/22/2021"
output: 
  html_document:
    toc: true 
    toc_float: true 
    toc_depth: 6
    code_folding: "hide"
---

We are going to run through the code base with just a couple variables in a model - in a slightly simplified workflow

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE)

library(tidyverse)
library(sf)
library(RSocrata) #opendata fn
library(viridis)
library(spatstat)
library(raster)
library(spdep)
library(FNN)
library(grid)
library(gridExtra)
library(knitr)
library(kableExtra)
library(tidycensus)
# functions
root.dir = "https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/DATA/"
source("https://raw.githubusercontent.com/urbanSpatial/Public-Policy-Analytics-Landing/master/functions.r")

```

## Read in Data from Chicago

This uses the Socrata package for some data sets.

Note where we bring in burglary data - you will want to vary this part to do your homework!

```{r}
policeDistricts <- 
  st_read("https://data.cityofchicago.org/api/geospatial/fthy-xz3r?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = dist_num)
  
policeBeats <- 
  st_read("https://data.cityofchicago.org/api/geospatial/aerh-rz74?method=export&format=GeoJSON") %>%
  st_transform('ESRI:102271') %>%
  dplyr::select(District = beat_num)

bothPoliceUnits <- rbind(mutate(policeDistricts, Legend = "Police Districts"), 
                         mutate(policeBeats, Legend = "Police Beats"))
robbery <- 
  read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2017/d62x-nvdr") %>%
    filter(Primary.Type == "ROBBERY") %>%
    mutate(x = gsub("[()]", "", Location)) %>%
    separate(x,into= c("Y","X"), sep=",") %>%
    mutate(X = as.numeric(X),Y = as.numeric(Y)) %>% 
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant")%>%
    st_transform('ESRI:102271') %>% 
    distinct()
    
#robbery$Description%>%unique()

chicagoBoundary <- 
  st_read(file.path(root.dir,"/Chapter5/chicagoBoundary.geojson")) %>%
  st_transform('ESRI:102271') 

#### RISK or NOT RISK

    ## Transit 

Lstops <- read_csv("https://raw.githubusercontent.com/bri-ne/MUSA_508_Lab/main/CTA_-_System_Information_-_List_of__L__Stops.csv")%>%
  mutate(x = gsub("[()]", "", Location)) %>%
  separate(x,into= c("Lat","Long"), sep=",")

Lstops.sf<- Lstops%>% st_as_sf(.,coords=c("Long", "Lat"),crs=4326)%>%st_transform('ESRI:102271')

busstops <- st_read("C:\\Users\\brine\\Documents\\GitHub\\MUSA-508\\MUSA_508_Lab_2021_10_26\\CTA_BusStops\\CTA_BusStops.shp")%>%
  st_transform('ESRI:102271')

    ## FARMERS MARKET
farmymarket <- read_csv("https://raw.githubusercontent.com/bri-ne/MUSA_508_Lab/main/Farmers_Markets_-_Map.csv")%>%st_as_sf(.,coords=c("LONGITUDE", "LATITUDE"),crs=4326)%>%st_transform('ESRI:102271')

    ## LIQUOR STORE
liquor<- read_csv("https://raw.githubusercontent.com/bri-ne/MUSA_508_Lab/main/Business_Licenses_-_Current_Liquor_and_Public_Places_of_Amusement_Licenses.csv")%>% drop_na(32:33)%>%st_as_sf(.,coords=c("LONGITUDE", "LATITUDE"),crs=4326)%>%st_transform('ESRI:102271')


    
```

## visualizing point data

Plotting point data and density

> How do we analyze point data?
>
> Are there other geometries useful to represent point locations?

```{r fig.width=6, fig.height=4}
# uses grid.arrange to organize indpendent plots
grid.arrange(ncol=2,
ggplot() + 
  geom_sf(data = chicagoBoundary) +
  geom_sf(data = robbery, colour="red", size=0.1, show.legend = "point") +
  labs(title= "Robberies, Chicago - 2017") +
  mapTheme(title_size = 14),

ggplot() + 
  geom_sf(data = chicagoBoundary, fill = "grey40") +
  stat_density2d(data = data.frame(st_coordinates(robbery)), 
                 aes(X, Y, fill = ..level.., alpha = ..level..),
                 size = 0.01, bins = 40, geom = 'polygon') +
  scale_fill_viridis() +
  scale_alpha(range = c(0.00, 0.35), guide = FALSE) +
  labs(title = "Density of Robberies") +
  mapTheme(title_size = 14) + theme(legend.position = "none"))
```

## Creating a fishnet grid

> What is a fishnet grid?

The `{sf}` package offers really easy way to create fishnet grids.

Examine the fishnet - the unique ID is crucial to building a data set!

```{r}
## using {sf} to create the grid
## Note the `.[chicagoBoundary] %>% ` line. This is needed to clip the grid to our data
fishnet <- 
  st_make_grid(chicagoBoundary,
               cellsize = 500, 
               square = TRUE) %>%
  .[chicagoBoundary] %>%            # <- MDH Added
  st_sf() %>%
  mutate(uniqueID = rownames(.))
```

### Aggregate points to the fishnet

> How can we aggregate points into a fishnet grid?

```{r}
## add a value of 1 to each crime, sum them with aggregate
## maybe 24 neighs?  no\
## approx a kfold like 24 folds 
crime_net <- 
  dplyr::select(robbery) %>% 
  mutate(countRobberies = 1) %>% 
 aggregate(., fishnet, sum) %>%
  mutate(countRobberies = replace_na(countRobberies, 0),
         uniqueID = rownames(.),
         cvID = sample(round(nrow(fishnet) / 24), 
                       size=nrow(fishnet), replace = TRUE))

ggplot() +
  geom_sf(data = crime_net, aes(fill = countRobberies), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of Roberies for the fishnet") +
  mapTheme()

# For demo. requires updated mapview package
# xx <- mapview::mapview(crime_net, zcol = "countBurglaries")
# yy <- mapview::mapview(mutate(burglaries, ID = seq(1:n())))
# xx + yy
```

## Modeling Spatial Features

> What features would be helpful in predicting the location of burglaries?
>
> What might these features be problematic?
>
> hint: for all the reasons we learned in class

```{r}
## only pulling a single variable for our model to keep it simple
## using Socrata again
abandonCars <- 
  read.socrata("https://data.cityofchicago.org/Service-Requests/311-Service-Requests-Abandoned-Vehicles/3c9v-pnva") %>%
    mutate(year = substr(creation_date,1,4)) %>% filter(year == "2017") %>%
    dplyr::select(Y = latitude, X = longitude) %>%
    na.omit() %>%
    st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "Abandoned_Cars")%>%dplyr::select(Legend)

liquor.fish <- liquor %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "liquor_licenses")%>%dplyr::select(Legend)

Lstops.sf <- Lstops.sf %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "L_stops")%>%dplyr::select(Legend)


busstops.sf <- busstops %>%
    st_transform(st_crs(fishnet)) %>%
    mutate(Legend = "bus_stops")%>%dplyr::select(Legend)





## Neighborhoods to use in LOOCV in a bit
neighborhoods <- 
  st_read("https://raw.githubusercontent.com/blackmad/neighborhoods/master/chicago.geojson") %>%
  st_transform(st_crs(fishnet)) 

```

#### How we aggregate a feature to our fishnet

This is an important chunk of code with some unfamiliar lines. We'll step through it.

```{r}

vars_net <- 
  rbind(abandonCars, liquor.fish,busstops.sf, Lstops.sf) %>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
    full_join(fishnet) %>%
    spread(Legend, count, fill=0) %>%
    st_sf() %>%
    dplyr::select(-`<NA>`) %>%
    na.omit() %>%
    ungroup()

liquor.vars_net <- liquor.fish%>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
  full_join(fishnet, by = "uniqueID") %>%
  spread(Legend, count, fill=0) %>%
  st_sf() %>%
  dplyr::select(-`<NA>`) %>%
  na.omit() %>%
  ungroup()


bus<- busstops.sf%>%dplyr::select( OBJECTID, Legend)%>%rename(STOP_ID=OBJECTID)
transit <- Lstops.sf%>%dplyr::select(STOP_ID, Legend)%>%rbind(., bus)
  

transit.vars_net <- transit%>%
  st_join(., fishnet, join=st_within) %>%
  st_drop_geometry() %>%
  group_by(uniqueID, Legend) %>%
  summarize(count = n()) %>%
  full_join(fishnet, by = "uniqueID") %>%
  spread(Legend, count, fill=0) %>%
  st_sf() %>%
  dplyr::select(-`<NA>`) %>%
  na.omit() %>%
  ungroup()

## 

# vars_net <- abandonCars %>%
#   spatially join abandonCars points to the fishnet polygon they are within %>%
#   drop the geometry attribute %>%
#   group_by each cells ID and the name of the feature %>%
#   summarize count the number of each point per grid cell %>%
#   join that summary back to spatial fishnet by cell ID %>%
#   "spread" from long to wide format and make column of our point count %>%
#   tell R that this should be an sf object %>%
#   remove a fussy column that appears b/c of NA %>%
#   get rid of rows with an NA in any column %>%
#   remove grouping so you are not tripped up later
```

## Nearest Neighbor Feature

> Review: what is NN and what does `k` represent in this function?

```{r}
# convinience to reduce length of function names.
st_c    <- st_coordinates
st_coid <- st_centroid

## create NN from abandoned cars
vars_net <- vars_net %>%
    mutate(Abandoned_Cars.nn = nn_function(st_c(st_coid(vars_net)), 
                                           st_c(abandonCars),
                                           k = 5))

transit.vars_net <- transit.vars_net %>%
    mutate(transit.nn = nn_function(st_c(st_coid(transit.vars_net)), 
                                           st_c(transit),
                                           k = 5))

liquor.vars_net <- liquor.vars_net %>%
    mutate(liquor.nn = nn_function(st_c(st_coid(liquor.vars_net)), 
                                           st_c(liquor.fish),
                                           k = 5))
```

> What changes if we make `k` a different number?

```{r}
## Visualize the NN feature
vars_net.long.nn <- 
  dplyr::select(vars_net, ends_with(".nn")) %>%
    gather(Variable, value, -geometry)




transit.vars_net.long.nn <- 
  dplyr::select(transit.vars_net, ends_with(".nn")) %>%
    gather(Variable, value, -geometry)


liquor.vars_net.long.nn <- 
  dplyr::select(liquor.vars_net, ends_with(".nn")) %>%
    gather(Variable, value, -geometry)

ggplot() +
      geom_sf(data = vars_net.long.nn, aes(fill=value), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Abandoned Car NN Distance") +
      mapTheme()




grid.arrange(ncol=2,
ggplot() +
      geom_sf(data = liquor.vars_net.long.nn, aes(fill=value), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Liquor Licenses NN Distance") +
      mapTheme(),

ggplot() +
      geom_sf(data = transit.vars_net.long.nn, aes(fill=value), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Transit Stops NN Distance") +
      mapTheme()
)




```

## Join NN feature to our fishnet

Since the counts were aggregated to each cell by `uniqueID` we can use that to join the counts to the fishnet.

```{r}
## important to drop the geometry from joining features

final_net <-
  left_join(crime_net, st_drop_geometry(vars_net), by="uniqueID") 





liquor.final_net <-
  left_join(crime_net, st_drop_geometry(liquor.vars_net), by="uniqueID") 
transit.final_net <-
  left_join(crime_net, st_drop_geometry(transit.vars_net), by="uniqueID") 

```

### Join in areal data

Using spatial joins to join *centroids* of fishnets to polygon for neighborhoods and districts.

> What issues arise when we try to join polygons to polygons in space?

```{r}

final_net <-
  st_centroid(final_net) %>%
    st_join(dplyr::select(neighborhoods, name), by = "uniqueID") %>%
    st_join(dplyr::select(policeDistricts, District), by = "uniqueID") %>%
      st_drop_geometry() %>%
      left_join(dplyr::select(final_net, geometry, uniqueID)) %>%
      st_sf() %>%
  na.omit()





liquor.final_net <-
  st_centroid(liquor.final_net) %>%
    st_join(dplyr::select(neighborhoods, name), by = "uniqueID") %>%
    st_join(dplyr::select(policeDistricts, District), by = "uniqueID") %>%
      st_drop_geometry() %>%
      left_join(dplyr::select(liquor.final_net, geometry, uniqueID)) %>%
      st_sf() %>%
  na.omit()


transit.final_net <-
  st_centroid(transit.final_net) %>%
    st_join(dplyr::select(neighborhoods, name), by = "uniqueID") %>%
    st_join(dplyr::select(policeDistricts, District), by = "uniqueID") %>%
      st_drop_geometry() %>%
      left_join(dplyr::select(transit.final_net, geometry, uniqueID)) %>%
      st_sf() %>%
  na.omit()

# for live demo
# mapview::mapview(final_net, zcol = "District")
```

## Local Moran's I for fishnet grid cells (CHANGES)

using {spdep} package to to build neighborhood weights and list to calculate local Moran's I.

Note that the code here is *different* than in the book - it has been updated to keep up with changes in packages.

> What is the difference between local and global Moran's I?

A little in depth version of the chunk below can be found:

Mendez C. (2020). Spatial autocorrelation analysis in R. R Studio/RPubs. Available at <https://rpubs.com/quarcs-lab/spatial-autocorrelation>

```{r}
## generates warnings from PROJ issues
## {spdep} to make polygon to neighborhoods... 
final_net.nb <- poly2nb(as_Spatial(final_net), queen=TRUE)

liquor.final_net.nb <- poly2nb(as_Spatial(liquor.final_net), queen=TRUE)
transit.final_net.nb <- poly2nb(as_Spatial(transit.final_net), queen=TRUE)

## ... and neighborhoods to list of weigths

final_net.weights <- nb2listw(final_net.nb, style="W", zero.policy=TRUE)


liquor.final_net.weights <- nb2listw(liquor.final_net.nb , style="W", zero.policy=TRUE)
transit.final_net.weights <- nb2listw(transit.final_net.nb, style="W", zero.policy=TRUE)

# print(final_net.weights, zero.policy=TRUE)
```

```{r}
## see ?localmoran
local_morans <- localmoran(final_net$Abandoned_Cars, final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()




liquor.local_morans <- localmoran(liquor.final_net$liquor_licenses, liquor.final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()

transit.final_net$stops <- transit.final_net$bus_stops + transit.final_net$L_stops

transit.local_morans <- localmoran(transit.final_net$stops, transit.final_net.weights, zero.policy=TRUE) %>% 
  as.data.frame()

# join local Moran's I results to fishnet


final_net.localMorans <- 
  cbind(local_morans, as.data.frame(final_net)) %>% 
  st_sf() %>%
  dplyr::select(Abandoned_Cars_Count = Abandoned_Cars, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)










liquor.final_net.localMorans <- 
  cbind(liquor.local_morans, as.data.frame(liquor.final_net)) %>% 
  st_sf() %>%
  dplyr::select(Liquor_Licenses_Count = liquor_licenses, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)


transit.final_net.localMorans <- 
  cbind(transit.local_morans, as.data.frame(transit.final_net)) %>% 
  st_sf() %>%
  dplyr::select(Transit_Stops_Count = stops, 
                Local_Morans_I = Ii, 
                P_Value = `Pr(z != E(Ii))`) %>%
  mutate(Significant_Hotspots = ifelse(P_Value <= 0.001, 1, 0)) %>%
  gather(Variable, Value, -geometry)


```

### Plotting local Moran's I results

This is a complex code chunk - it's a loop which builds ggplots of local Moran's for each of your `vars`

> What does a significant hot spot tell us about the distribution of burglaries?

```{r}
## This is just for plotting
vars <- unique(final_net.localMorans$Variable)
varList <- list()


liquor.vars <- unique(liquor.final_net.localMorans$Variable)
liquor.varList <- list()

transit.vars <- unique(transit.final_net.localMorans$Variable)
transit.varList <- list()





for(i in vars){
  varList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(final_net.localMorans, Variable == i), 
              aes(fill = Value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme(title_size = 14) + theme(legend.position="bottom")
  }

do.call(grid.arrange,c(varList, ncol = 4, top = "Local Morans I statistics, Burglary"))



liquor.final_net.localMorans

#### liquor plot

for(i in liquor.vars){
  liquor.varList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(liquor.final_net.localMorans, Variable == i), 
              aes(fill = Value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme(title_size = 14) + theme(legend.position="bottom")
  }

do.call(grid.arrange,c(liquor.varList, ncol = 4, top = "Local Morans I statistics, Liquor Licenses"))


#### transit plot 

for(i in transit.vars){
  transit.varList[[i]] <- 
    ggplot() +
      geom_sf(data = filter(transit.final_net.localMorans, Variable == i), 
              aes(fill = Value), colour=NA) +
      scale_fill_viridis(name="") +
      labs(title=i) +
      mapTheme(title_size = 14) + theme(legend.position="bottom")
  }

do.call(grid.arrange,c(transit.varList, ncol = 4, top = "Local Morans I statistics, Transit Stops"))






```

## Distance to Hot spot

Using NN distance to a hot spot location

```{r}
# generates warning from NN

bothvars.final_net <- liquor.final_net%>%cbind(transit.final_net)



final_net <- final_net %>% 
  mutate(abandoned.isSig = 
           ifelse(local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(abandoned.isSig.dist = 
           nn_function(st_c(st_coid(final_net)),
                       st_c(st_coid(filter(final_net, 
                                           abandoned.isSig == 1))), 
                       k = 1))


#### separate

liquor.final_net <- liquor.final_net %>% 
  mutate(robbery.isSig = 
           ifelse(liquor.local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(robbery.isSig.dist = 
           nn_function(st_c(st_coid(liquor.final_net)),
                       st_c(st_coid(filter(liquor.final_net, 
                                           robbery.isSig == 1))), 
                       k = 1))


transit.final_net <- transit.final_net%>% 
  mutate(robbery.isSig = 
           ifelse(transit.local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(robbery.isSig.dist = 
           nn_function(st_c(st_coid(transit.final_net)),
                       st_c(st_coid(filter(transit.final_net, 
                                           robbery.isSig == 1))), 
                       k = 1))


#### together

bothvars.final_net <- bothvars.final_net %>% 
  mutate(liquor.isSig = 
           ifelse(liquor.local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(liquor.isSig.dist = 
           nn_function(st_c(st_coid(liquor.final_net)),
                       st_c(st_coid(filter(liquor.final_net, 
                                           liquor.isSig == 1))), 
                       k = 1)) %>% 
  mutate(transit.isSig = 
           ifelse(transit.local_morans[,5] <= 0.001, 1, 0)) %>%
  mutate(transit.isSig.dist = 
           nn_function(st_c(st_coid(transit.final_net)),
                       st_c(st_coid(filter(transit.final_net, 
                                           transit.isSig == 1))), 
                       k = 1))


## What does k = 1 represent?
```

> What does `k = 1` above mean in terms of measuring nearest neighbors?

### Plot NN distance to hot spot

```{r}
ggplot() +
      geom_sf(data = final_net, aes(fill=abandoned.isSig.dist), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Abandoned Car NN Distance") +
      mapTheme()









grid.arrange(ncol=2,
ggplot() +
      geom_sf(data = liquor.final_net, aes(fill=liquor.isSig.dist), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Liquor Licenses NN Distance") +
      mapTheme(),
ggplot() +
      geom_sf(data = transit.final_net, aes(fill=transit.isSig.dist), colour=NA) +
      scale_fill_viridis(name="NN Distance") +
      labs(title="Transit Stops NN Distance") +
      mapTheme()

)





```


## Correlation 

```{r}

#### for counts
correlation.long <-
  st_drop_geometry(final_net) %>%
    dplyr::select(-uniqueID, -cvID,  -name, -District) %>%
    gather(Variable, Value, -countRobberies)



 

 ##### for nn

correlation.cor <-
  correlation.long %>%
    group_by(Variable) %>%
    summarize(correlation = cor(Value, countRobberies, use = "complete.obs"))
 
 
 ##### PLOTS
  
ggplot(correlation.long, aes(Value, countRobberies)) +
  geom_point(size = 0.1) +
  geom_text(data = correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Robberies count as a function of risk factors") +
  plotTheme() 
 
### liquor plot    
ggplot(liquor.correlation.long, aes(Value, countRobberies)) +
  geom_point(size = 0.1) +
  geom_text(data = liquor.correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Robberies count as a function of risk factors") +
  plotTheme()
 
 
 
 
 ### transit plot    
ggplot(transit.correlation.long, aes(Value, countRobberies)) +
  geom_point(size = 0.1) +
  geom_text(data = transit.correlation.cor, aes(label = paste("r =", round(correlation, 2))),
            x=-Inf, y=Inf, vjust = 1.5, hjust = -.1) +
  geom_smooth(method = "lm", se = FALSE, colour = "black") +
  facet_wrap(~Variable, ncol = 2, scales = "free") +
  labs(title = "Robberies count as a function of risk factors") +
  plotTheme()
```



## Modeling and CV

Leave One Group Out CV on spatial features

```{r results='hide'}

# View(crossValidate)

## define the variables we want
reg.ss.vars <- c("Abandoned_Cars.nn", "abandoned.isSig.dist")

## RUN REGRESSIONS
reg.ss.spatialCV <- crossValidate(
  dataset = final_net,
  id = "name",                           
  dependentVariable = "countBurglaries",
  indVariables = reg.ss.vars) %>%
    dplyr::select(cvID = name, countBurglaries, Prediction, geometry)
```

```{r}
# calculate errors by NEIGHBORHOOD
error_by_reg_and_fold <- 
  reg.ss.spatialCV %>%
    group_by(cvID) %>% 
    summarize(Mean_Error = mean(Prediction - countBurglaries, na.rm = T),
              MAE = mean(abs(Mean_Error), na.rm = T),
              SD_MAE = mean(abs(Mean_Error), na.rm = T)) %>%
  ungroup()

error_by_reg_and_fold %>% 
  arrange(desc(MAE))
error_by_reg_and_fold %>% 
  arrange(MAE)

## plot histogram of OOF (out of fold) errors
error_by_reg_and_fold %>%
  ggplot(aes(MAE)) + 
    geom_histogram(bins = 30, colour="black", fill = "#FDE725FF") +
  scale_x_continuous(breaks = seq(0, 11, by = 1)) + 
    labs(title="Distribution of MAE", subtitle = "LOGO-CV",
         x="Mean Absolute Error", y="Count") 
```

## Density vs predictions (CHANGES)

The `spatstat` function gets us kernal density estimates with varying search radii.

Note that the code here is *different* than in the book - it has been updated to keep up with changes in packages.

```{r}
# demo of kernel width
burg_ppp <- as.ppp(st_coordinates(burglaries), W = st_bbox(final_net))
burg_KD.1000 <- spatstat.core::density.ppp(burg_ppp, 1000)
burg_KD.1500 <- spatstat.core::density.ppp(burg_ppp, 1500)
burg_KD.2000 <- spatstat.core::density.ppp(burg_ppp, 2000)
burg_KD.df <- rbind(
  mutate(data.frame(rasterToPoints(mask(raster(burg_KD.1000), as(neighborhoods, 'Spatial')))), Legend = "1000 Ft."),
  mutate(data.frame(rasterToPoints(mask(raster(burg_KD.1500), as(neighborhoods, 'Spatial')))), Legend = "1500 Ft."),
  mutate(data.frame(rasterToPoints(mask(raster(burg_KD.2000), as(neighborhoods, 'Spatial')))), Legend = "2000 Ft.")) 

burg_KD.df$Legend <- factor(burg_KD.df$Legend, levels = c("1000 Ft.", "1500 Ft.", "2000 Ft."))

ggplot(data=burg_KD.df, aes(x=x, y=y)) +
  geom_raster(aes(fill=layer)) + 
  facet_wrap(~Legend) +
  coord_sf(crs=st_crs(final_net)) + 
  scale_fill_viridis(name="Density") +
  labs(title = "Kernel density with 3 different search radii") +
  mapTheme(title_size = 14)
```

```{r}

as.data.frame(burg_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) %>%
   ggplot() +
     geom_sf(aes(fill=value)) +
     geom_sf(data = sample_n(burglaries, 1000), size = .5) +
     scale_fill_viridis(name = "Density") +
     labs(title = "Kernel density of 2017 burglaries") +
     mapTheme(title_size = 14)
```

## Get 2018 crime data

Let's see how our model performed relative to KD on the following year's data.

```{r}
burglaries18 <- 
  read.socrata("https://data.cityofchicago.org/Public-Safety/Crimes-2018/3i3m-jwuy") %>% 
  filter(Primary.Type == "BURGLARY" & 
         Description == "FORCIBLE ENTRY") %>%
  mutate(x = gsub("[()]", "", Location)) %>%
  separate(x,into= c("Y","X"), sep=",") %>%
  mutate(X = as.numeric(X),
         Y = as.numeric(Y)) %>% 
  na.omit %>%
  st_as_sf(coords = c("X", "Y"), crs = 4326, agr = "constant") %>%
  st_transform('ESRI:102271') %>% 
  distinct() %>%
  .[fishnet,]
```

```{r}
burg_KDE_sf <- as.data.frame(burg_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) %>%
  mutate(label = "Kernel Density",
         Risk_Category = ntile(value, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
           Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
           Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
           Risk_Category >= 1 & Risk_Category  <= 29 ~ "1% to 29%")) %>%
  cbind(
    aggregate(
      dplyr::select(burglaries18) %>% mutate(burgCount = 1), ., sum) %>%
    mutate(burgCount = replace_na(burgCount, 0))) %>%
  dplyr::select(label, Risk_Category, burgCount)
```

Note that this is different from the book, where we pull a model out of a list of models we've created. For your homework, you'll be creating multiple models.

```{r}
burg_risk_sf <-
  reg.ss.spatialCV %>%
  mutate(label = "Risk Predictions",
         Risk_Category = ntile(Prediction, 100),
         Risk_Category = case_when(
         Risk_Category >= 90 ~ "90% to 100%",
         Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
         Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
         Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
         Risk_Category >= 1 & Risk_Category <= 29 ~ "1% to 29%")) %>%
  cbind(
    aggregate(
      dplyr::select(burglaries18) %>% mutate(burgCount = 1), ., sum) %>%
      mutate(burgCount = replace_na(burgCount, 0))) %>%
  dplyr::select(label,Risk_Category, burgCount)
```

We don't do quite as well because we don't have very many features, but still pretty good.

```{r}
rbind(burg_KDE_sf, burg_risk_sf) %>%
  na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category, -geometry) %>%
  ggplot() +
    geom_sf(aes(fill = Risk_Category), colour = NA) +
    geom_sf(data = sample_n(burglaries18, 3000), size = .5, colour = "black") +
    facet_wrap(~label, ) +
    scale_fill_viridis(discrete = TRUE) +
    labs(title="Comparison of Kernel Density and Risk Predictions",
         subtitle="2017 burglar risk predictions; 2018 burglaries") +
    mapTheme(title_size = 14)
```

```{r}
rbind(burg_KDE_sf, burg_risk_sf) %>%
  st_set_geometry(NULL) %>% na.omit() %>%
  gather(Variable, Value, -label, -Risk_Category) %>%
  group_by(label, Risk_Category) %>%
  summarize(countBurglaries = sum(Value)) %>%
  ungroup() %>%
  group_by(label) %>%
  mutate(Rate_of_test_set_crimes = countBurglaries / sum(countBurglaries)) %>%
    ggplot(aes(Risk_Category,Rate_of_test_set_crimes)) +
      geom_bar(aes(fill=label), position="dodge", stat="identity") +
      scale_fill_viridis(discrete = TRUE) +
      labs(title = "Risk prediction vs. Kernel density, 2018 burglaries") +
      theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
```






```{r}


rob_risk_2sf <-
  filter(reg.summary, Regression == "Spatial LOGO-CV: Spatial Process")# %>%
  mutate(label = "Risk Predictions",
         Risk_Category = ntile(Prediction, 100),
         Risk_Category = case_when(
           Risk_Category >= 90 ~ "90% to 100%",
           Risk_Category >= 70 & Risk_Category <= 89 ~ "70% to 89%",
           Risk_Category >= 50 & Risk_Category <= 69 ~ "50% to 69%",
           Risk_Category >= 30 & Risk_Category <= 49 ~ "30% to 49%",
           Risk_Category >= 1 & Risk_Category <= 29 ~ "1% to 29%"))


 #aggregate(, ., sum)
      B <-dplyr::select(robbery18) %>% mutate(robCount = 1) %>%
      mutate(robCount = replace_na(robCount, 0)) #%>%
  dplyr::select(label,Risk_Category, robCount)


sum(rob_risk_sf$countRobberies) #11790

sum(B$robCount) #9673

B2<- aggregate(B, rob_risk_sf, sum)





burg_KDE_sf <- as.data.frame(burg_KD.1000) %>%
  st_as_sf(coords = c("x", "y"), crs = st_crs(final_net)) %>%
  aggregate(., final_net, mean) 





grid.arrange(ncol=3,

ggplot() +
    geom_sf(data= reg.ss.Kfold.spatialCV, aes(fill = Prediction))+
    scale_fill_viridis() +
    labs(title = "Robbery : K-fold") +
    mapTheme() + theme(legend.position="bottom"),


ggplot() +
    geom_sf(data= reg.ss.LOGO.spatialCV, aes(fill = Prediction))+
    scale_fill_viridis() +
    labs(title = "Robbery : LOGO-CVn") +
    mapTheme() + theme(legend.position="bottom"),


ggplot() +
  geom_sf(data = crime_net, aes(fill = countRobberies), color = NA) +
  scale_fill_viridis() +
  labs(title = "Count of Roberies for the fishnet") +
  mapTheme()+ theme(legend.position="bottom"))









```


